---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# dplyr-cli

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
<!-- badges: end -->

`dplyr-cli` uses [{littler}](https://cran.r-project.org/package=littler) to 
run dplyr commands on CSV files in the terminal.

`dplyr-cli` makes use of the terminal pipe `|` instead of the magrittr pipe (`%>%`)
to run sequences of commands.

```
cat mtcars.csv | group_by cyl | summarise "mpg = mean(mpg)" | kable
#> | cyl|      mpg|
#> |---:|--------:|
#> |   4| 26.66364|
#> |   6| 19.74286|
#> |   8| 15.10000|
```

## Motivation

I wanted to be able to do quick hacks on CSV files on the command line without
actually starting a proper R session.


## What dplyr commands are supported?

Any command of the form `dplyr::verb(.data, code)`

Currently two extra commands are supported which are not part of `dplyr`.

* `csv` performs no dplyr command, but only outputs the input data as CSV to stdout
* `kable` performs no dplyr command, but only outputs the input data as a
  `knitr::kable()` formatted string to stdout


## Limitations

* When using special shell characters such as `()`, you'll have to quote 
  your code arguments
* Every command runs in a separate R session - startup overhead can get 
  prohibitive.

## Usage

```{sh}
./dplyr.sh --help
```



## Installation

`dplyr-cli` is run from the shell but at every invocation is starting a new 
rsession where the following packages are expected to be installed:


```{r eval=FALSE}
install.packages('littler')  # Scripting front-end for R
install.packages('glue')     # text manipulation
install.packages('dplyr')    # data manipulation
install.packages('docopt')   # CLI description language
```

You'll then need to download the shell script from my github and put `dplyr.sh`
somewhere in your path.

```
git clone https://github.com/coolbutuseless/dplyr-cli
```


# Example data

```{r}
write.csv(mtcars, "mtcars.csv", row.names = FALSE)
```

# Example 1 - Basic Usage

```{sh}
# cat contents of input CSV into dplyr-cli.  
# Use '-c' to output CSV if this is the final step
cat mtcars.csv | ./dplyr.sh filter -c mpg == 21

# Put quotes around any commands which contain special characters like <>()
cat mtcars.csv | ./dplyr.sh filter -c "mpg < 11"

# Combine dplyr commands with shell 'head' command
./dplyr.sh select --file mtcars.csv -c cyl | head -n 6
```


# Example 2 - Simple piping of commands (with shell pipe, not magrittr pipe)

```{sh}
cat mtcars.csv | \
   ./dplyr.sh mutate "cyl2 = 2 * cyl"  | \
   ./dplyr.sh filter "cyl == 8" | \
   ./dplyr.sh kable
```


# Example 3 - set up some aliases for convenience


```{sh}
alias mutate="./dplyr.sh mutate"
alias filter="./dplyr.sh filter"
alias select="./dplyr.sh select"
alias summarise="./dplyr.sh summarise"
alias group_by="./dplyr.sh group_by"
alias ungroup="./dplyr.sh ungroup"
alias count="./dplyr.sh count"
alias arrange="./dplyr.sh arrange"
alias kable="./dplyr.sh kable"


cat mtcars.csv | group_by cyl | summarise "mpg = mean(mpg)" | kable
```


## Security warning

`dplyr-cli` uses `eval(parse(text = ...))` on user input.  Do not expose this 
program to the internet or random users under any circumstances.


## Inspirations

* [xsv](https://github.com/BurntSushi/xsv) - a fast CSV command line toolkit 
  written in Rust
* [jq](https://stedolan.github.io/jq/) - a command line JSON processor.
